---
title: "coursera 8"
author: "Muskaan"
date: "21/11/2020"
output: md_document
---
TITLE: Prediction Assignment by Muskaan Parmar

SUMMARY: The goal of this assignment is to use the data of 6 participants obtained from accelerometers on the belt, forearm, arm, and dumbell and predict the manner in which they exercise. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. Since, people regularly quantify how much of a particular activity they do, but they rarely quantify how well they do it which would be dealt in this assignment. We will build a model, use cross validation and thus make choices. Also, we will use the model on 20 different test cases.

1. Setting the directory and loading the data
```{r , echo=TRUE}
setwd("~/R/Coursera 8")
library(randomForest)
library(lattice)
library(ggplot2)
library(doParallel)
library(rpart)
library(caret)   
library(rpart.plot)

train <- read.csv("pml-training.csv", na.strings=c("NA","","#DIV/0!"))
test <- read.csv("pml-testing.csv", na.strings=c("NA", "", "#DIV/0!"))
#loading data and removing NA,#DIV/0! and blank values from data
#head(train)
print("After removing NA,#DIV/0! and blank values from the two datasets:")
print("Dimension of training data=")
dim(train)
print("Dimension of testing data=")
dim(test)
```

2.Data pre-processing
```{r , echo=TRUE}
train2<-train[,-c(1:7)]
test2 <-test[,-c(1:7)]
#removing index, timestamp, new window, num window and subject name i.e. first 7 columns
print("After removing non predictors from the two datasets:")
print("Dimension of training data=")
dim(train2)
print("Dimension of testing data=")
dim(test2)
#checking for non zero values in training dataset
train3<-nzv(train2[,-ncol(train2)],saveMetrics=TRUE)
#row(train3)
```

3.Partitioning train data into validation(testing) set and training set
```{r , echo=TRUE}
intr<- createDataPartition(train2$classe, p = 0.6, list = FALSE)
training<- train2[intr,]#training set(60%)
validation<- train2[-intr,]#validation set(40%)
print("After partitioning training data into validation set(40%) and training set(60%) :")
print("Dimension of training set=")
dim(training)
print("Dimension of validation set=")
dim(validation)
```

4.Model building by cross validation using Random Forest algorithm
```{r , echo=TRUE}

mfn <- "myModel.RData"
if (!file.exists(mfn)) 
{
  nc <- makeCluster(detectCores() - 1)
  registerDoParallel(cores=nc)
  getDoParWorkers() # 3    
  
  myModel  <- train(classe ~ ., data = training, method = "rf", metric = "Accuracy",         preProcess=c("center", "scale"), trControl=trainControl(method = "cv", number = 4, p= 0.60, allowParallel = TRUE ) 
  )
  save(myModel , file = "myModel.RData")
  stopCluster(nc)
}else 
{
  load(file = mfn, verbose = TRUE)
}
print(myModel, digits=4)
```

5.Prediction
```{r , echo=TRUE}
predTest <- predict(myModel, newdata=validation)
```

6.Confusion Matrix
```{r , echo=TRUE}

confusionMatrix(predTest, factor(validation$classe))
```
The out of sample error is 0.0037.
The accuracy is 0.9964 and lies within the 95% confidence interval.


7.Complete data about the model
```{r , echo=TRUE}
myModel$finalModel
varImp(myModel)
```

8. Quiz Coursera
The testing is now performed on the Quiz set.
```{r , echo=TRUE}
print(predict(myModel, newdata=test2))
```
The above sequence is the set of answers obtained for the Quiz.


CITATIONS: The data for this project comes from : http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har

